Finding Quality Issues in SKOS Vocabularies

### 2012-06-04 analysis.tex
more todos for cr version
changes according to antoine's comments

\subsection{Results}

The results of this analysis are summarized in Table~\ref{tab:results}, which shows the absolute number of possibly affected resources for each quality checking function and vocabulary combination. Numbers marked with an asterisk (*) were obtained by extrapolating from subsets containing 5\% of the respective vocabulary resources.

@@

% LABELING AND DOCUMENTATION ISSUES
We found labeling and documentation issues in all vocabularies.
% Omitted or Invalid Language Tags
\texttt{MeSH}, \texttt{PXV}, \texttt{Pressinfo}, and \texttt{LVAk} omit language tags with their labeling properties, \texttt{LCSH} with the \texttt{skos:note} property. \texttt{STW} does not use language tags with 2 instances of \texttt{skos:definition}.
% Incomplete Language Coverage
\texttt{AGROVOC} covers 25 languages but no single concept is labeled in all languages, in \texttt{Meketre} all concepts have English but only some of them French labels assigned. \texttt{STW}, which is expressed mainly in English and German, has many concepts with incomplete language coverage because it (i) links to non-authoritative concepts that are only labeled in German and (ii) uses the private, but valid language tag \texttt{x-other} with some of its concept labels.
%The reasons for the high number of concepts with incomplete language coverage in \texttt{STW} are that many concepts (i) link to non-authoritative concepts that are only labeld in German and (ii) lack labels tagged with the private, but valid language tag \texttt{x-other} that few concepts pull in. 
% Undocumented Concepts
\texttt{Geonames}, which defines a concept scheme of ``feature codes'', is the only vocabulary in our dataset, which has at least one documentation property assigned to all of its concepts. All other vocabularies have a significant number of undocumented concepts.
% Label Conflicts
We also detected possible label conflicts in half of the vocabularies. \texttt{PXV}, for instance, uses the string ``primary peroxisomal enzyme deficiency'' with two concepts in the same concept scheme, but once with a \texttt{skos:prefLabel} and another time with a \texttt{skos:altLabel} property. In \texttt{NAICS} we couldn't detect any labeling issues but found that it expresses statements with \texttt{skosxl:prefLabel} as predicate and plain literals as object, which contradicts the SKOS-XL specification.

@@

The difference between the number of concepts and the number of authoritative concepts in Table~\ref{tab:vocabs} already indicates which vocabularies are linked with other SKOS vocabularies.
% Missing In-Links
However, except \texttt{NYTP} and \texttt{Geonames}, no vocabulary has a high number of estimated in-links from other web resources.
%Estimation of these numbers is rather inaccurate since currently we only rely on the Sindice index. 
% Missing Out-Links
Also the number of out-links is rather low: \texttt{NYTP}, \texttt{IPSV}, and \texttt{STW} are the three exceptions, which are fully linked to other Web resources. The one concept with missing out-links in \texttt{NAICS} is the object of a \texttt{skos:broaderTransitive} relation. One reason for a high number of missing out-links could also be that the links were not available in the main thesaurus file, which is at least the case for \texttt{AGROVOC}.
% Broken Links
Even though we could not determine the exact number of broken links because of the large number of links to resolve  (over 400K in \texttt{Eurovoc}, over 500K in \texttt{LCSH}), we found that broken links are a common issue in most vocabularies.
% Undefined SKOS Resources
Undefined SKOS resources seem to be a minor issue, because we could only find two of them in all vocabularies: \texttt{MeSH} introduces \texttt{skos:annotation} and \texttt{IPSV} uses the deprecated \texttt{skos:prefSymbol} property.

>>>>

\subsection{Results and Discussion}

The results of this analysis are summarized in Table~\ref{tab:results}, which shows the absolute number of possibly affected resources for each quality checking function and vocabulary. Numbers marked with an asterisk (*) were obtained by extrapolating from subsets containing 5\% of the respective vocabulary resources.

@@

% LABELING AND DOCUMENTATION ISSUES
We found labeling and documentation issues in all vocabularies.
% Omitted or Invalid Language Tags
\texttt{MeSH}, \texttt{PXV}, \texttt{Pressinfo}, and \texttt{LVAk} omit language tags with their labeling properties, \texttt{LCSH} with the \texttt{skos:note} property. \texttt{STW} does not use language tags with 2 instances of \texttt{skos:definition}.
% Incomplete Language Coverage
\texttt{AGROVOC} covers 25 languages but no single concept is labeled in all languages, in \texttt{Meketre} all concepts have English but only some of them French labels assigned. \texttt{STW}, which is expressed mainly in English and German, has many concepts with incomplete language coverage because it (i) links to non-authoritative concepts that are only labeled in German and (ii) uses the private, but valid language tag \texttt{x-other} with some of its concept labels.
%The reasons for the high number of concepts with incomplete language coverage in \texttt{STW} are that many concepts (i) link to non-authoritative concepts that are only labeld in German and (ii) lack labels tagged with the private, but valid language tag \texttt{x-other} that few concepts pull in. 
% Undocumented Concepts
\texttt{Geonames}, which defines a concept scheme of ``feature codes'', is the only vocabulary in our dataset, which has at least one documentation property assigned to all of its concepts. All other vocabularies have a significant number of undocumented concepts.
% Label Conflicts
We also detected possible label conflicts in half of the vocabularies. \texttt{PXV}, for instance, uses the string ``primary peroxisomal enzyme deficiency'' with two concepts in the same concept scheme, but once with a \texttt{skos:prefLabel} and another time with a \texttt{skos:altLabel} property. In \texttt{NAICS} we could not detect any labeling issues but found that it expresses statements with \texttt{skosxl:prefLabel} as predicate and plain literals as object, which contradicts the SKOS-XL specification.

@@

The difference between the number of concepts and the number of authoritative concepts in Table~\ref{tab:vocabs} already indicates which vocabularies are linked with other SKOS vocabularies.
% Missing In-Links
However, except \texttt{NYTP} and \texttt{Geonames}, no vocabulary has a high number of estimated in-links from other web resources.
%Estimation of these numbers is rather inaccurate since currently we only rely on the Sindice index. 
% Missing Out-Links
Also the number of out-links is rather low: \texttt{NYTP}, \texttt{IPSV}, and \texttt{STW} are the three exceptions, which are fully linked to other Web resources. The one concept with missing out-links in \texttt{NAICS} is the object of a \texttt{skos:broaderTransitive} relation. One reason for a high number of missing out-links is that the links were not available in the main thesaurus file, which is at least the case for \texttt{AGROVOC}.
% Broken Links
Even though we could not determine the exact number of broken links because of the large number of links to resolve  (over 400K in \texttt{Eurovoc}, over 500K in \texttt{LCSH}), we found that broken links are a common issue in most vocabularies.
% Undefined SKOS Resources
Undefined SKOS resources seem to be a minor issue, because we could only find two of them in all vocabularies: \texttt{MeSH} introduces \texttt{skos:annotation} and \texttt{IPSV} uses the deprecated \texttt{skos:prefSymbol} property.

### 2012-06-04 background.tex
changes according to antoine's comments

% Data quality in general
The problem of ``vocabulary quality'' is closely related to the more general one of ``data quality`'' and has been discussed in data and information systems research (cf.~\cite{Batini2009}). Pipino et al.~\cite{Pipino2002} argue that dealing with data quality should involve both ``subjective perceptions of the individuals'' and ``objective measurements based on the data''. We see our work as a contribution to the latter.

@@

Typical applications of controlled vocabularies are classification, indexing, auto-completion, query reformulation, or serving as a glossary. As we discussed in detail in earlier work~\cite{Nagy2011}, these areas impose specific requirements on vocabulary features, such as structure, availability, and documentation. Quality aspects of controlled vocabularies have already been discussed in standardized guidelines~\cite{ISO25964-1:2011,Z39.19:2005}, manuals~\cite{Svenonius2003,Hedden2010,Aitchison2000,Harpring2010}, tutorials~\cite{Soergel2002}, and scholarly articles \cite{Coronado2009,Kless2010}. These most often rely on manual, precise analysis of individual statements in the data, as in~\cite{spero2008}. Our work builds on this literature, but focuses on the less intellectually loaded checks, which can be automatized to assist vocabulary users or publishers.

>>>>

% Data quality in general
The problem of ``vocabulary quality'' is closely related to the more general one of ``data quality'' and has been discussed in data and information systems research (cf.~\cite{Batini2009}). Pipino et al.~\cite{Pipino2002} argue that dealing with data quality should involve both ``subjective perceptions of the individuals'' and ``objective measurements based on the data''. We see our work as a contribution to the latter.

@@

Typical applications of controlled vocabularies are classification, indexing, auto-completion, query reformulation, or serving as a glossary. As we discussed in detail in earlier work~\cite{Nagy2011}, these areas impose specific requirements on vocabulary features, such as structure, availability, and documentation. Quality aspects of controlled vocabularies have already been discussed in standardized guidelines~\cite{ISO25964-1:2011,Z39.19:2005}, manuals~\cite{Aitchison2000,Harpring2010,Hedden2010,Svenonius2003}, tutorials~\cite{Soergel2002}, and scholarly articles \cite{Coronado2009,Kless2010}. These most often rely on manual, precise analysis of individual statements in the data, as in~\cite{spero2008}. Our work builds on this literature, but focuses on the less intellectually loaded checks, which can be automatized to assist vocabulary users or publishers.

### 2012-06-04 criteria.tex
changes according to antoine's comments

% Origin and rationale
SKOS defines a set of properties that link resources with RDF Literals, which are plain text natural language strings with an optional language tag. This includes the labeling properties {\texttt{rdfs:label}, \texttt{skos:prefLabel}, \texttt{skos:altLabel}, \texttt{skos:hiddenLabel} and also SKOS documentation properties, such as \texttt{skos:note} and subproperties thereof. Literals should be tagged consistently~\cite{Vrandecic2010}, because omitting language tags or using non-standardized, private language tags in a SKOS vocabulary could unintentionally limit the result set of language-dependent queries.
% Function defintion
A SKOS vocabulary can be checked for omitted and invalid language tags by iterating over all resources in $IR$ and finding those that have labeling or documentation property relations to plain literals in $LV$ with missing or invalid language tags, i.e., tags that are not defined in RFC3066\footnote{RFC 3066: Tags for the Identification of Languages \url{http://www.ietf.org/rfc/rfc3066.txt}}.

@@

% Origin and rationale
The SKOS Primer~\cite{Isaac2009} recommends that ``no two concepts have the same preferred lexical label in a given language when they belong to the same concept scheme''. This issue could affect application scenarios such as auto-completion, which proposes labels based on user input. Although these extra cases are acceptable for some thesauri, we generalize the above recommendation and search for all concept pairs with their respective \texttt{skos:prefLabel}, \texttt{skos:altLabel} or \texttt{skos:hiddenLabel} property values meeting a certain similarity threshold defined by a function $sim:LV \times LV \rightarrow [0,1]$.
% Function defintion
The default, built-in similarity function checks for case-insensitive string equality with a threshold equal to 1. Label conflicts can be found by iterating over all concept pairs $AC \times AC$, applying $sim$ to every possible label combination, and collecting those pairs with at least one label combination meeting or exceeding a specified similarity threshold. We handle this issue under the Closed World Assumption, because data on concept scheme membership may lack and concepts may be linked to concepts with similar labels in other vocabularies.
% Since data on concept scheme membership may lack because of the Open World Assumption, we handle this issue under the Closed World Assumption.

@@

% Origin and rationale
When vocabularies are being published on the Web, SKOS concepts become linkable resources. Estimating the number of in-links and identifying the concepts without any in-links, can indicate the importance of a concept.
% Function definition
We estimate the number of in-links by iterating over all elements in $AC$ and querying the Sindice\footnote{\url{http://sindice.com/} indexes the Web of Data, which is composed of pages with semantic markup in RDF, RDFa, Microformats, or Microdata. Currently it covers approximately 230M documents with over 11 billion triples.} SPARQL endpoint for triples containing the concept's URI in the object part. Empty query results are indicators for missing in-links.

@@

% Origin and rationale
The SKOS model is defined within the namespace \url{http://www.w3.org/2004/02/skos/core#}. However, some vocabularies use resources from within this namespace, which are unresolvable for two main reasons: vocabulary creators ``invented'' new terms within the SKOS namespace instead of introducing them in a separate namespace, or they use ``deprecated'' SKOS elements like \texttt{skos:subject}.
% Function definition
Undefined SKOS resources can be identified by iterating over all resources in $IR$ and returning those resources (i) that are contained in the list of deprecated resources\footnote{See \url{http://www.w3.org/TR/skos-reference/#namespace}} or (ii) are identified by a URI in the SKOS namespace but are not defined in the current version of the SKOS ontology.

>>>>

% Origin and rationale
SKOS defines a set of properties that link resources with RDF Literals, which are plain text natural language strings with an optional language tag. This includes the labeling properties {\texttt{rdfs:label}, \texttt{skos:prefLabel}, \texttt{skos:altLabel}, \texttt{skos:hiddenLabel} and also SKOS documentation properties, such as \texttt{skos:note} and subproperties thereof. Literals should be tagged consistently~\cite{Vrandecic2010}, because omitting language tags or using non-standardized, private language tags in a SKOS vocabulary could unintentionally limit the result set of language-dependent queries.
% Function defintion
A SKOS vocabulary can be checked for omitted and invalid language tags by iterating over all resources in $IR$ and finding those that have labeling or documentation property relations to plain literals in $LV$ with missing or invalid language tags, i.e., tags that are not defined in RFC3066\footnote{Tags for the Identification of Languages \url{http://www.ietf.org/rfc/rfc3066.txt}}.

@@

% Origin and rationale
The SKOS Primer~\cite{Isaac2009} recommends that ``no two concepts have the same preferred lexical label in a given language when they belong to the same concept scheme''. This issue could affect application scenarios such as auto-completion, which proposes labels based on user input. Although these extra cases are acceptable for some thesauri, we generalize the above recommendation and search for all concept pairs with their respective \texttt{skos:prefLabel}, \texttt{skos:altLabel} or \texttt{skos:hiddenLabel} property values meeting a certain similarity threshold defined by a function $sim:LV \times LV \rightarrow [0,1]$.
% Function defintion
The default, built-in similarity function checks for case-insensitive string equality with a threshold equal to 1. Label conflicts can be found by iterating over all (authoritative) concept pairs $AC \times AC$, applying $sim$ to every possible label combination, and collecting those pairs with at least one label combination meeting or exceeding a specified similarity threshold. We handle this issue under the Closed World Assumption, because data on concept scheme membership may lack and concepts may be linked to concepts with similar labels in other vocabularies.
% Since data on concept scheme membership may lack because of the Open World Assumption, we handle this issue under the Closed World Assumption.

@@

% Origin and rationale
When vocabularies are published on the Web, SKOS concepts become linkable resources. Estimating the number of in-links and identifying the concepts without any in-links, can indicate the importance of a concept.
% Function definition
We estimate the number of in-links by iterating over all elements in $AC$ and querying the Sindice\footnote{\url{http://sindice.com/} indexes the Web of Data, which is composed of pages with semantic markup in RDF, RDFa, Microformats, or Microdata. Currently it covers approximately 230M documents with over 11 billion triples.} SPARQL endpoint for triples containing the concept's URI in the object part. Empty query results are indicators for missing in-links.

@@

% Origin and rationale
The SKOS model is defined within the namespace \url{http://www.w3.org/2004/02/skos/core#}. However, some vocabularies use resources from within this namespace, which are unresolvable for two main reasons: vocabulary creators ``invented'' new terms within the SKOS namespace instead of introducing them in a separate namespace, or they use ``deprecated'' SKOS elements like \texttt{skos:subject}.
% Function definition
Undefined SKOS resources can be identified by iterating over all resources in $IR$ and returning those (i) that are contained in the list of deprecated resources\footnote{See \url{http://www.w3.org/TR/skos-reference/#namespace}} or (ii) are identified by a URI in the SKOS namespace but are not defined in the current version of the SKOS ontology.

### 2012-06-04 introduction.tex
changes according to antoine's comments

% What is SKOS?
The Simple Knowledge Organization System (SKOS)~\cite{SkosReference2008} is a standard model for sharing and linking controlled vocabularies (thesauri, classification systems, etc.) on the Web. Organizations like, e.g., the European Union\footnote{EuroVoc, \url{http://eurovoc.europa.eu/}}, the United Nations\footnote{AGROVOC, \url{http://aims.fao.org/website/AGROVOC-Thesaurus/sub}}, or the UK government\footnote{Integrated Public Sector Vocabulary (IPSV), \url{http://doc.esd.org.uk/IPSV}} publish SKOS representations of their vocabularies on the Web so that they can easily be accessed by humans and machines.

@@

We aim at contributing to the ongoing community efforts to bridge that gap between model-level integrity constraints and domain-specific quality aspects. Our goal is to help taxonomists in identifying possible quality issues in SKOS vocabularies and to give them a set of computable quality checking functions that, in combination with the taxonomists' experience and domain expertise, can serve as quality indicators for vocabularies. Finding such quality issues also gives important feedback on the overall vocabulary design process and should, at the end, lead to better vocabularies. With defining computable quality checking functions we tackle the problem of quality assessment from an objective perspective. Subjective perception of quality is not within the scope of this work. The main contributions of this work can be summarized as follows:

>>>>

% What is SKOS?
The Simple Knowledge Organization System (SKOS)~\cite{SkosReference2008} is a standard model for sharing and linking controlled vocabularies (thesauri, classification systems, etc.) on the Web. Organizations like, e.g., the European Union\footnote{EuroVoc, \url{http://eurovoc.europa.eu/}}, the United Nations\footnote{AGROVOC, \url{http://aims.fao.org/standards/agrovoc/about}}, or the UK government\footnote{Integrated Public Sector Vocabulary (IPSV), \url{http://doc.esd.org.uk/IPSV}} publish SKOS representations of their vocabularies on the Web so that they can easily be accessed by humans and machines.

@@

We aim at contributing to the ongoing community efforts to bridge that gap between model-level integrity constraints and domain-specific quality aspects. Our goal is to help taxonomists in identifying possible quality issues in SKOS vocabularies and to give them a set of computable quality checking functions that, in combination with the taxonomists' experience and domain expertise, can serve as quality indicators for vocabularies. Finding such quality issues also gives important feedback on the overall vocabulary design process and should, in the end, lead to better vocabularies. With defining computable quality checking functions we tackle the problem of quality assessment from an objective perspective. Subjective perception of quality is not within the scope of this work. Our contribution can be summarized as follows:
